---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# prsTools

<!-- badges: start -->
<!-- badges: end -->

The goal of prsTools is to ...

## Installation
The prsTools R package is installed by using the install_github command


```{r}
library(devtools)
# install_github("ararder/prsTools")

```


## Getting the metrics
The most common (and basic) form of PRS analysis is to use the PRS to predict an outcome y,
in a regression framework. In R, the most common use is to use the glm function.

In many cases, one wants to fit many different variations of the same model, with different covariates PRS:es.

glm_prs_metrics provided a framework for doing this programmatically, and for extracting relevant metrics.


```{r example}
library(prsTools)
library(tidyverse)

test_df <- tibble(
  disease_status = sample(c(0,1), 10, replace=TRUE),
  prs_1 = rnorm(10),
  prs_2 = rnorm(10),
  pc1 = rnorm(10),
  age = sample(seq(30,60), 10, replace=TRUE),
)
```

The base formula is used for calculating R2 in logistic regression.
and is used as a baseline comparison

```{r }
base_formula <- as.formula("disease_status ~ pc1")
full_formula <- as.formula("disease_status ~ prs_1 + pc1")

# get commonly used metrics to evaluate the PRS. The model defaults to returning the odds-ratio/beta for the first predictor
# in this case prs_1
glm_prs_metrics(df = test_df, base_formula = base_formula, formula = full_formula)
```

For case-control phenotypes, we need to put the R2 measurement on the liabiltiy scale
to be able to compare against cohorts with different case/control ratios
for this, we need to populatiom prevalence of the disease. here i put it to 20%

```{r }
glm_prs_metrics(df = test_df, base_formula = base_formula, formula = full_formula, pop_prev = 0.2)
```


we can return odds-ratio by setting the argument odds_ratio = TRUE instead of the beta
this is helpful for generating confidence intervals on the odds-ratio scale
```{r }
glm_prs_metrics(df = test_df, base_formula = base_formula, formula = full_formula, pop_prev = 0.2, odds_ratio = TRUE)
```


Often you want to run lots of possible variations of this model.

```{r}
formula_list <- list(
  full_formula,
  as.formula("disease_status ~ prs_1 + prs_2 + pc1"),
  as.formula("disease_status ~ prs_1 + prs_2 + pc1 + age"),
  as.formula("disease_status ~ prs_2 + pc1 + age")
)

# it's now programmatically easy to compare these different models.
map_df(formula_list, glm_prs_metrics, df = test_df, base_formula = base_formula, pop_prev = 0.2)

```


# Commands for Slurm
TBA - helper functions for generating slurm.sh files inside R

```{r pressure, echo = FALSE}
# -A [accountname] - for uppmax
# -t [12:00:00] hour/minutes/seconds
# -o |output of slurm file]
# -p [core/node]
# -n [number of cores]
# -J [name of the job]
```


